---
title: 'DataSpace Report'
output:
  html_document
---

# HVTN 505 Highlights from: _Antibody Fc effector functions and IgG3 associate with decreased HIV-1 risk_ (Neidich 2019)

The intention of this report is to demonstrate how to reproduce
research using integrated and non-integrated datasets from DataSpace.
DataSpace has always provided integrated data to its users, that is,
data that has been integrated into our standardized schema, but 
DataSpace also offers study-specific datasets and asssays to
suppliment data that provided in the app in an integrated
format.

Here, using DataSpaceR’s access to both integrated and non-integrated
datasets, we can reproduce some of the results from the published
paper, Antibody Fc effector functions and IgG3 associate with
decreased HIV-1 risk by (Neidich 2019).

[https://www.ncbi.nlm.nih.gov/pubmed/31589165](https://www.ncbi.nlm.nih.gov/pubmed/31589165)

This analysis is not a complete reproduction of the original work,
focusing only on a subset of the final analysis.  The data was not
processed precisely the same way, although we do attempt to recreate
similar processing in effort to match the original analysis.

A few of the key results were reproduced below. Parts of the result
section of the original paper are quoted below and appear in order as
they do in the original paper to highlight those findings:

> The response rate for HIV-1 Env Con S gp140 FcγRIIa was 98.7%

> In a regression analysis controlling for age, race, behavior risk,
> and body mass index (BMI), the primary hypothesis outcomes of IgG
> ADCP and FcγRIIa responses to HIV-1 Env glycoprotein Con S gp140
> were both inversely correlated with HIV-1 acquisition risk.

> HIV-1 Env IgG3 correlates with decreased HIV-1 risk in a human HIV-1
> vaccine efficacy trial.

> Vaccine-matched gp140 IgG3 response rates were high, ranging from
> 55% to 90%...

> V1V2 IgG3 response rates were very low (<20%) and were not included
> in the test for association with HIV-1 acquisition.

<link rel="stylesheet" type="text/css" href="../../_webdav/CAVD/@files/CAVD-DataSpace-Reports/dependencies/css/learn_report_margins.css">

## Reading in integrated and non-integrated data

We will use a combination of `DataSpaceR` and `Rlabkey` to retrieve
the integrated and non-integrated datasets respectively.

```{r set-options, echo=FALSE}

knitr::opts_chunk$set(cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      highlight = TRUE,
                      fig.align = "center",
                      fig.width = 11)

```

```{r retrieve-process}

library(data.table)
library(DataSpaceR)
library(Rlabkey)
library(ggplot2)

# connect and review available datasets

con <- connectDS()
stu <- con$getStudy("vtn505")
stu$availableDatasets

## load integrated data with DataSpaceR

bam <- copy(stu$getDataset("BAMA"))
dem <- copy(stu$getDataset("Demographics"))

## load non-integrated data with DataSpaceR

fca <- copy(stu$getDataset("Fc Array"))
adc <- copy(stu$getDataset("ADCP"))
dms <- copy(stu$getDataset("DEM SUPP"))

dem <- dem[,`:=`(
    trt = vapply(study_arm, function(x) switch(x, "C1" = 0, "T1" = 1), c(1))
    )];

dem <- merge(dem, dms[, .(ParticipantId, bhvrisk, HIVwk28, cc_cohort, cc_strata)], by = "ParticipantId");
setnames(dem, c("age_enrollment", "bmi_enrollment"), c("age", "bmi"))

## Antigen name mapping coming soon to DataSpace. This mapping was
## done manually for this work.

antigenNameMapping <- function(x){
    switch(tolower(make.names(x)), 
           "a1.con.env03.140.cf"      = "A1 Con Env03 gp140 CF",
           "ae.a244.v1v2.tags.293f"   = "AE A244 V1V2 Tags/293F",
           "b.con.env03.140.cf"       = "B Con Env03 gp140 CF",
           "b.mn.v3.gp70"             = "B MN V3 gp70",
           "c.1086.v2.tags.293f"      = "C 1086 V2 Tags/293F",
           "c.1086c_v1_v2.tags"       = "C 1086C V1V2 Tags",
           "c.con.env03.140.cf_avi"   = "C Con Env03 gp140 CF avi",
           "c.con.env03.140.cf"       = "C Con Env03 gp140 CF",
           "con.6.gp120.b"            = "Con 6 gp120/B",
           "con.s.gp140.cfi"          = "Con S gp140 CFI",
           "gp41"                     = "gp41",
           "gp70.v1v2..a."            = "gp70 V1V2 (A)",
           "gp70_b.casea2.v1.v2.169k" = "gp70 B CaseA2 V1/V2/169K",
           "gp70_b.casea_v1_v2"       = "gp70 B CaseA V1V2",
           "gp70_c.1086c.v1.v2.293f"  = "gp70 C 1086C V1/V2/293F",
           "p24"                      = "p24",
           "vrc.a_avi"                = "VRC A gp140 avi",
           "vrc.b.gp140"              = "VRC B gp140",
           "vrc.c_avi"                = "VRC C gp140 avi",
           "vrc_a_gp70v1v2_avi"       = "VRC A gp70 V1V2 avi",
           x)
}

```

## Findings

### _The response rate for HIV-1 Env Con S gp140 FcγRIIa was 98.7%_

```{r}

fcaResp <- copy(fca)

fcaResp <- fcaResp[, antigen:=vapply(antigen, antigenNameMapping, c("a"))]

fcaResp <- dem[,.(ParticipantId, trt)
               ][fcaResp, on = "ParticipantId"]

fcaResp[study_day == 196 & antibody_isotype == "FcgRIIa H131" & antigen == "Con S gp140 CFI" & trt == 1,
    round(sum(fca_response)/length(fca_response), 3)
]

```

### _In a regression analysis controlling for age, race, behavior risk,and body mass index (BMI), the primary hypothesis outcomes of IgG ADCP and FcγRIIa responses to HIV-1 Env glycoprotein Con S gp140 were both inversely correlated with HIV-1 acquisition risk._ 

The type of regression used in the original analysis was a two phase
regression model provided by `osDesign::tps()`.

```{r}

library(osDesign)

assignStratumNumber <- function(stn){
    switch(stn,
           "Vacc/Blk_Hisp/[18.4, 25)" = 1,
           "Vacc/Blk_Hisp/[25, 29.8)" = 2,
           "Vacc/Blk_Hisp/[29.8, 40)" = 3,
           "Vacc/White/[18.4, 25)"    = 4,   
           "Vacc/White/[25, 29.8)"    = 5,
           "Vacc/White/[29.8, 40)"    = 6,
           -1 
           )
}

dem <- dem[,stratum_num := vapply(cc_strata, assignStratumNumber, c(1))]

## define case and controls counts for tps

dem <- dem[,`:=`(
    nn0 = sum(HIVwk28 == 0),
    nn1 = sum(HIVwk28 == 1)
), by = stratum_num]

## process data: log transform fca, subset to day 196 and RIIa, cast on antigen

fca <- fca[, lfi:=log(1+pmax(0, mfi_delta))]
fcc <- dcast(fca[antibody_isotype == "FcgRIIa H131" & study_day == 196,], ParticipantId~antigen, value.var="lfi")
setnames(fcc, names(fcc)[-1], paste0("R2a.", names(fcc)[-1]))

dat <- adc[study_day == 196, .(ParticipantId, avg_phagocytosis_score)
           ][fcc, on = "ParticipantId"]
setnames(dat, "avg_phagocytosis_score", "ADCP")
setnames(dat, names(dat), make.names(names(dat)))

## reduce data to treatment subjects only

dat <- dem[dat
           ][trt == 1]

## make stratum group counts

nnx <- dem[,.N, by = .(stratum_num, nn0, nn1)
           ][order(stratum_num)
             ][stratum_num > 0,.(nn0, nn1)]

## fit tps regression

tpsFit <- function(var, dat){
    reg_form <- paste("HIVwk28 ~ age + race + bhvrisk + bmi + ", var)

    dat <- dat[!is.na(dat[[var]])]
    dat[[var]] <- scale(dat[[var]])

    fit <- tps(
        formula = reg_form,
        data    = dat,
        nn0     = nnx$nn0,
        nn1     = nnx$nn1,
        group   = dat$stratum_num
    )

    res <- summary(fit)$coef
    res <- cbind(
        res[,c("Value","Emp p")],
        "lower bound" = res[,"Value"] - ( res[, "Emp SE"] * 1.96 ),
        "upper bound" = res[,"Value"] + ( res[, "Emp SE"] * 1.96 )
    )
    res[,c("Value", "lower bound", "upper bound")] <- exp(res[,c("Value", "lower bound", "upper bound")])
    colnames(res) <- c("OR", "p.value", "lower", "upper")
    res <- data.table(data.frame(res)[var,])[,intercept:=var]
    setcolorder(res, c(ncol(res), 1:(ncol(res)-1)))
}

results <- rbindlist(lapply(c("ADCP", grep("^R2a", names(dat), value = TRUE)), tpsFit, dat = dat))
setnames(results, "intercept", "variable")

results <- results[, variable := sapply(gsub("^R2a\\.", "", variable), antigenNameMapping)]

knitr::kable(
         results[order(p.value)],
         caption = "Results from univariant regression of a two phase design.",
         col.names = c("Variable", "OR", "p value", "lower", "upper"),
         digits = 3
       )

## plot OR and mfi delta highlighting very low p values

dat <- dat[,grep("^R2a|ADCP", names(dat)), with = F]
dat <- melt(
    dat,
    value.name = "log_mfi_delta",
    variable.name = "variable"
)

dat <- dat[, variable:=sapply(gsub("^R2a\\.", "", variable), antigenNameMapping)]

dat <- results[dat[, mfi_delta:=exp(log_mfi_delta)
                   ][, med_mfi_delta:=median(mfi_delta, na.rm = TRUE), by = variable],
               on = "variable"
                 ][, sig:=ifelse(p.value <= 0.005, "p ≤ 0.005", "p > 0.005")]

ggplot(data = dat[,.N, by = .(med_mfi_delta, OR, variable, sig)],
       aes(med_mfi_delta, OR, color = sig)) +
    geom_point() +
    geom_text(aes(label=variable), hjust=0, vjust=0, show.legend = F) +
    theme_bw() + 
    scale_color_manual(name = "p value",
                       values = c("p ≤ 0.005" = "black", "p > 0.005" = "grey70")) +
    ggtitle("Odds Ratio of HIV-1 Risk and Median MFI Delta") +
    xlab("Median MFI Delta") +
    ylab("Odds Ratio")

```

### _HIV-1 Env IgG3 correlates with decreased HIV-1 risk in a human HIV-1 vaccine efficacy trial._

```{r}

bam <- copy(stu$getDataset("BAMA"))

bam <- bam[, lfi:=log(1+pmax(0, mfi_delta))]

bam <- dcast(
    bam[antibody_isotype == "IgG3" & visit_day == 196,],
    ParticipantId~antigen,
    value.var = "lfi"
)

setnames(bam, names(bam),     make.names(names(bam)))
setnames(bam, names(bam)[-1], paste0("mfi.", names(bam)[-1]))

## reset stratum_num

dem <- dem[,stratum_num := vapply(cc_strata, assignStratumNumber, c(1))]

## define case and controls counts for tps

dem <- dem[,`:=`(
    nn0 = sum(HIVwk28 == 0),
    nn1 = sum(HIVwk28 == 1)
), by = stratum_num]

bam <- dem[,.(ParticipantId, stratum_num, trt, HIVwk28,  age, race, bhvrisk, bmi, nn0, nn1)
           ][bam
             ][trt == 1]

results <- rbindlist(lapply(grep("^mfi.", names(bam), value = TRUE), tpsFit, dat = bam))
setnames(results, "intercept", "antigen")
results <- results[, antigen_name:=sapply(gsub("^mfi\\.", "", antigen), antigenNameMapping)]

knitr::kable(results[order(p.value), .(antigen_name, OR, p.value, lower, upper)],
             caption = "IgG3 Binding Antibody Correlated to HIV-1 Risk of Infection",
             col.names = c("antigen", "OR", "p value", "lower", "upper"),
             digits = 3)

mnames <- grepl("^mfi", names(bam))

bam <- results[,antigen_name:=NULL][
    melt(
        bam,
        names(bam)[!mnames],
        names(bam)[mnames],
        value.name = "log_mfi_delta",
        variable.name = "antigen"
    ), on = "antigen"]

bam <- bam[, antigen:=sapply(gsub("^mfi\\.", "", antigen), antigenNameMapping)]
bam <- bam[, mfi_delta:=exp(log_mfi_delta)]
bam <- bam[, antigen:=paste0(antigen, "\n(OR = ", round(OR, 2), ")")]

clr <- colorRampPalette(c("royalblue4", "steelblue3", "grey70", "grey90"))(10)

ggplot(bam[!is.na(mfi_delta)], aes(antigen, mfi_delta)) +
    geom_boxplot(aes(fill = p.value)) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.margin = margin(0.5, 0.5, 0.5, 1, "cm")) +
    scale_x_discrete(limits = unique(bam$antigen)) +
    scale_y_continuous(trans='log10') +
    scale_fill_gradientn(name = "p value", colors = clr, trans='log10') +
    ggtitle("IgG3 Binding Antibody MFI Delta Correlated to HIV-1 risk") +
    xlab("Antigen (Odds Ratio)") +
    ylab("MFI Delta")

```

### _Vaccine-matched gp140 IgG3 response rates were high, ranging from 55% to 90%..._

```{r}

bam <- copy(stu$getDataset("BAMA"))
bam <- bam[, antigen:=sapply(make.names(antigen), antigenNameMapping)]

bam <- dem[,.(ParticipantId, stratum_num, trt, HIVwk28,  age, race, bhvrisk, bmi, nn0, nn1)
           ][bam[visit_day == 196]
             ][trt == 1 & antibody_isotype == "IgG3"
               ][order(antigen)]

knitr::kable(
           bam[
               grepl("140", antigen),
               .(response_rate = sum(response_call == 1)/length(response_call)),
               by = antigen
           ],
           col.names = c("Antigen", "Response Rate"),
           digits = 3)

```

### _V1V2 IgG3 response rates were very low (<20%) and were not included in the test for association with HIV-1 acquisition._

```{r}

knitr::kable(
           bam[
               grepl("V1|V2", antigen),
               .(response_rate = sum(response_call == 1)/length(response_call)),
               by = antigen
           ],
           col.names = c("Antigen", "Response Rate"),
           digits = 3)

```

```{r}

ggplot(bam[,.(response_rate = sum(response_call == 1)/length(response_call)), by = antigen],
       aes(reorder(antigen, -response_rate), response_rate)) +
    geom_bar(stat="identity", ) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("BAMA Response Rate for IgG3 Biniding Antibodies") + 
    xlab("Antigen") +
    ylab("Response Rate")

```
