---
title: "DataSpace Report"
output: "html_document"
---

<link rel="stylesheet" type="text/css" href="../../_webdav/CAVD/@files/CAVD-DataSpace-Reports/dependencies/css/learn_report_margins.css">

```{r set-options, include=FALSE}

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  highlight = TRUE,
  tidy = FALSE,
  fig.align = "center",
  fig.width = 10)
 
```

```{r load-packages, include=FALSE}

library(DataSpaceR)
library(data.table)
library(ggplot2)

```

# Magnitude Breadth Curves Using DataSpaceR

The following report is intended to show a workflow for interfacing
with DataSpaceR's mAb utilities to create magnitude breadth curves.

For a more detailed explanation of DataSpaceR's mAb utilites, see the
DataSpaceR vignettes.

## Accessing and exploring NAb mAb data with DataSpaceR

First, connect to DataSpace, then display the summary of the mAb grid
via `mabGridSummary`. See other documentation associated with this
package for setting up proper credentials for accesing data using
DataSpaceR.

The `mabGridSummary` table will be a running summary of the results we
have in our current filter.

Showing only the top 10 rows below.

```{r using-mab-grid}

con <- connectDS()
con$mabGridSummary[1:10]

```

It is possible to query the grid as one would a `data.frame` type R
object. The mAb grid is actually stored as a `data.table` object.

The `mabGridSummary` field associated with the connection object is
where one can find the mAb grid. Shown below are a couple of examples
of how the grid can be queried.

```{r}

con$mabGridSummary[con$mabGridSummary$geometric_mean_curve_ic50 < 0.01]
con$mabGridSummary[hxb2_location %like% "gp160"]
con$mabGridSummary[mab_mixture %like% "PGT"]
 
```

Once the grid has been explored, we can select a list of mAbs we would
like data for.

In this case we are selecting all PGT mabs that exist in a mixture and
the individual components for those mixtures.

To filter the grid, use the `filterMabGrid` method on the connection
object. One can filter on any value found in any field in the
`mabGrid` table.
 
The API will throw a warning if there is a value that is passed that
is not found in the column that is filtered.

```{r select-mabs}

names(con$mabGrid)

pgtMix <- con$mabGridSummary[mab_mixture %like% "PGT" & mab_mixture %like% "\\+", mab_mixture]
pgtMix <- c(pgtMix, unique(unlist(strsplit(pgtMix, " \\+ |/"))))
sort(pgtMix)

con$filterMabGrid("mab_mixture", pgtMix)

```

View the grid by calling `mabGridSummary` or `mabGrid`, then get the
mAb object associated with the grid using the `getMab` method.

Using the mAb object, one can get the NAb mAb data for the mAbs that
were selected.

```{r view-then-extract}

con$mabGridSummary

mab <- con$getMab()
names(mab)

mab <- mab$nabMab

```

Review how many unique viruses are found in each protocol assayed for
each mAb.

```{r review-selection-prot}

vir <- unique(mab[,.(prot, mab_mix_name_std, clade, neutralization_tier, virus)])
dcast(vir, mab_mix_name_std~prot, value.var = "virus", fun.aggregate = length)

```

View the names of the first 50 viruses.

```{r review-selection-virus}

head(sort(unique(mab$virus)), 50)

```

Review the specific clades, tiers, and dilutions found in the
selection.

```{r review-selection-clade-tier}

dcast(vir, mab_mix_name_std~clade, value.var = "virus", fun.aggregate = length)
dcast(vir, mab_mix_name_std~neutralization_tier, value.var = "virus", fun.aggregate = length)

```

```{r review-selection-concentration}

dcast(mab, mab_mix_name_std~mab_concentration, value.var = "virus", fun.aggregate = length)

```

View virus counts for different groups of mabs.

```{r select-virus-panel1}

castTab <- dcast(mab, virus~mab_mix_name_std, value.var = "virus", fun.aggregate = length)
binGrp  <- apply(castTab[,-1], 1, function(x) paste0(as.numeric(x != 0), collapse = ""))
tabGrp  <- table(binGrp)

tabGrp

```

## Analyzing and visualizing NAb mAb data via magnitude breadth curves

We will analyze a set of viruses with the most mAb/virus combinations 
for the group of PGT combination and singleton mAbs queried above.

```{r select-virus-panel2}

members <- data.table(tabGrp, members = sapply(strsplit(names(tabGrp), ""), function(x) sum(as.numeric(x))))
members[,comboCnt:=members * N] 
selGrp <- members[order(comboCnt, decreasing = T), binGrp][1]

castTab <- castTab[binGrp == selGrp,]

mab <- mab[virus %in% castTab$virus]

unique(mab[,.(mab_mix_name_std, virus)])[,.(virus_count = .N), by=mab_mix_name_std]

```

To include as much data as possible, we will set censored titer
IC50 results (-Inf and +Inf) to the min and max concentrations
associated with those values. -Inf values are set to the minimum
concentration tested, and +Inf values are set to the maximum
concentration tested. This will assure that there are no missing
values when calculating magnitude.

Where multiple protocols test the same virus mAb combination (see
table above), the geometric mean of the titer IC50 results is
calculated. This will yield one titer IC50 per mAb virus combination.

```{r average-across-replicates}

mab <- mab[, `:=`(
    titer_curve_ic50_mod = 
        ifelse(titer_curve_ic50 == -Inf, min_concentration,
        ifelse(titer_curve_ic50 == Inf, max_concentration,
               titer_curve_ic50))
)]

mab <- mab[, `:=`(
  titer_curve_ic50_gmean = 10^mean(log10(titer_curve_ic50_mod))
), by = c("mab_mix_label", "virus")]

```

`magnitude` is the titer IC50 concentration. `breadth` is
the cumulative proportion of isolates having titer IC50 concentrations
at or below the concentraion on the x axis.

Plot magnitude as derived from IC50:

```{r plot-mb-ic50}

mb50 <- unique(mab[,.(mab_mix_label, titer_curve_ic50_gmean, virus)])
mb50 <- mb50[, magnitude := titer_curve_ic50_gmean]
mb50 <- mb50[, mag_order := as.numeric(factor(magnitude, ordered = T)), by = "mab_mix_label"]
mb50 <- mb50[, breadth   := mag_order/max(mag_order), by = "mab_mix_label"]

ggplot(mb50) +
    geom_line(aes(magnitude, breadth, color = mab_mix_label)) +
    theme_bw() +
    scale_x_log10() +
    ggtitle("Magnitude Breadth Curve using Titer IC50") +
    theme(legend.direction ="vertical")+
    guides(color=guide_legend(ncol=2))

```

Calculate the 1st, 2nd, and 3rd quartiles of magnitude for each mAb.

```{r get-quartiles}

qdt50 <- mb50[,
{
    qvals = quantile(magnitude)[c(2,3,4)]
    list(quart = qvals,
         qname = as.numeric(substr(names(qvals), 1, 2)))
}
, by = "mab_mix_label"]

qdt50Cast <- dcast(qdt50, mab_mix_label~qname, value.var = "quart")
qdt50Cast

```

### What mAbs have the lowest magnitude's for their first, second, and third quartiles?

```{r}

minMagQuart <- lapply(qdt50Cast[, -1, with = F], function(x) list(min(x), qdt50Cast$mab_mix_label[x == min(x)]))
minMagQuart

```

### Which virsues are not neutralized at any IC50 at or below our lowest 3rd quartile magnitude for any mAb?

```{r}

qVal <- minMagQuart[["75"]][[1]]

mbSel50 <- mb50[, .(magnitude=min(magnitude)), by=virus][magnitude >= qVal,]

virNeut <- unique(mbSel50[,virus])
virNeut

```

### What mAbs were tested against those viruses that showed lower IC50 results?

Search DataSpace for those identified viruses. We can use column names
from the `mabGrid` field to subset the mAb grid.

Select mAbs from grid where the magnitude of titer IC50 geometric mean
is less than the 3rd quartile magnitude of the previous mAb selection.

```{r}

con$resetMabGrid()

con$filterMabGrid("virus", virNeut)

gapMab <- con$mabGridSummary[!mab_mixture %in% unique(mab$mab_mix_label), mab_mixture]

con$filterMabGrid("mab_mixture", gapMab)

gapMab <- con$getMab()$nabMab

gapMab <- gapMab[, `:=`(
    titer_curve_ic50_mod = 
        ifelse(titer_curve_ic50 == -Inf, min_concentration,
        ifelse(titer_curve_ic50 == Inf, max_concentration,
               titer_curve_ic50))
)]

gapMab <- gapMab[, titer_curve_ic50_gmean:=10^mean(log10(titer_curve_ic50_mod)), by = "mab_mix_label"]
gapMab <- gapMab[titer_curve_ic50_gmean < qVal]

```

Show the titer IC50 values for those mAb and virus combinations.

```{r}

unique(gapMab[,.(mab_mix_label, titer_curve_ic50, virus)])

```

## Session information

```{r session-info}

sessionInfo()

```
