---
output: html_document
---

```{r include = F, echo = F}

library(data.table)

srvDir <- "/labkey/labkey/files"
webDir <- "../../_webdav"
depDir <- "/CAVD/@files/CAVD-DataSpace-Reports/dependencies/mab_visualizations"

setwd(paste0(srvDir, depDir))
source("./data_create.R")

```

```{r appendData, results = "asis", echo = F}

cat("<script>", appendDataObjects(mabDat), "</script>")

```

<script>

    (() => {

        let reportsdir = `r paste0(webDir, depDir)`;

        d3
            .select("head")
            .append("link")
            .attr("type", "text/css")
            .attr("rel", "stylesheet")
            .attr("href", reportsdir + "/css/mab_visualization_styles.css");
        
        [
            "/hm_functions.js",
            "/mb_functions.js",
            "/nc_functions.js",
            "/data_process.js",
            "/hm.js",
            "/mb.js",
            "/nc.js",
            "/neutralizationGrid.js"
        ].forEach(src => {
            var script = document.createElement('script');
            script.type = "text/javascript"
            script.src = reportsdir + src;
            script.async = false;
            document.head.appendChild(script);
        });

    })();

    (() => {
    
        let cover = document.createElement("div");
        cover.setAttribute("id", "mab_report_temp_cover");
        
        let styles = {
            'backgroundColor' : "#000000",
            'opacity'         : 0.6,
            'position'        : "absolute",
            'left'            : "-10%",
            'height'          : "100%",
            'width'           : "110%",
            'zIndex'          : 1000
        }
        
        Object.assign(cover.style, styles)
        document.querySelector("div[class='labkey-knitr'").prepend(cover)
        
        let removeCover = setInterval(function(){
            let rep = document.querySelectorAll(".mabreportcont");
            if(rep.length != 0 && this.mabReportData.plotStatus.length == 3){
                clearInterval(removeCover);
                [...rep].forEach(r => r.style.display = "none");
                document.querySelector("#mab_report_temp_cover").outerHTML = "";
            }
        },100);

    })();

    var mabActiveTab = function(e, tabname){
        var tabs = document.querySelectorAll(".mabreporttab");
        var cont = document.querySelectorAll(".mabreportcont");
        tabs.forEach(d => d.className = d.className.replace("active", ""));
        cont.forEach(d => d.style.display = "none");
        [...tabs].filter(d => d.id.includes(tabname)).forEach(d => d.className += " active");
        [...cont].filter(d => d.id.includes(tabname)).forEach(d => d.style.display = "block");
    }

</script>

<div class="mabreporttabs">
  <h1 class="mabreporttab" id="mab_neutcrv" onclick="mabActiveTab(event, 'mab_neutcrv')">Neutralization Curves</h1>
  <h1 class="mabreporttab" id="mab_neutgrd" onclick="mabActiveTab(event, 'mab_neutgrd')">Neutralization Grid</h1>
  <h1 class="mabreporttab" id="mab_mbcurve" onclick="mabActiveTab(event, 'mab_mbcurve')">Magnitude Breadth Curves</h1>
  <h1 class="mabreporttab" id="mab_heatmap" onclick="mabActiveTab(event, 'mab_heatmap')">Titer Heatmap</h1>
</div>

<div class="mabreportconts">
  <div class="mabreportcont" id="mab_neutcrv"></div>
  <div class="mabreportcont" id="mab_mbcurve"></div>
  <div class="mabreportcont" id="mab_heatmap"></div>
  <div class="mabreportcont" id="mab_neutgrd"></div>
</div>
